---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# earthquakesWithR

The goal of `earthquakesWithR` is to give a few handy visualization tools for the NOAA data set. It provides to the user of this package the following functions:  
- **geom_timeline()** - which is a specific geom for the `ggplot2` package  
- **geom_timeline_label()** - which is an extension for the above geom  
- **eq_location_clean()** - which is a helper function, cleaning the NOAA dataset  
- **eq_clean_data()** - a function which loads and applies a few cleaning function to it  
- **eq_map()** - a function creating a `leaflet` map and visualizing the earth quakes  
- **eq_create_label()** - is a helper function, creating a nice pop-ups in the leaflet map  

## Installation

You can install `earthquakesWithR` from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("moralmar/earthquakesWithR")
```

## Data

The Significant Earthquake Database contains information on destructive earthquakes from 2150 B.C. to the present that meet at least one of the following criteria: Moderate damage (approximately $1 million or more), 10 or more deaths, Magnitude 7.5 or greater, Modified Mercalli Intensity X or greater, or the earthquake generated a tsunami.  
Feel free to browse through the code as follows:  

```{r example}
# To load the earth quake data and have a look:
data("NOAAearthquakes")

# the following is a spatial polygon data frame, which merely gives a nice touch to the leaflet map
data("world")
```




