---
title: "Vignette - Package earthquakesWithR"
author: "mmorales"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette - Package earthquakesWithR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```


A brief overview was already given in the `README.md` file. Here in this vignette, we will examine futher exampmles and try to go to the limits of the package. This document is splitted in two main sections:  

- **Earth quakes and Timeline** - What can be done and its limits  
- **Earth quakes and Leaflet map** - What can be done and its limits

You can install `earthquakesWithR` from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("moralmar/earthquakesWithR")
```

## Earth quakes and Timeline

Let's investigate a few exmaple with the timeline functions give by the package

```{r, message=FALSE, warning=FALSE, fig.width = 7.2, fig.asp = .4}
require(earthquakesWithR)
require(magrittr) # for piping operator
require(dplyr) # for filter/subsetting functions
require(ggplot2) # should be obvious :)
require(ggthemes) # for the theme_classic()

CentralEurope <- c("AUSTRIA", "GERMANY", "HUNGARY", "LIECHTENSTEIN", "POLAND", "SLOVENIA", "SWITZERLAND")

# Load the data and filter it accordingly to your needs
data("NOAAearthquakes")
quakes_CentEU <- NOAAearthquakes %>%
        filter(COUNTRY %in% CentralEurope & YEAR >= 0)

# single country without labels
ggplot() +
        geom_timeline(data = quakes_CentEU, aes(x = date, y = COUNTRY, color = TOTAL_DEATHS, size = EQ_PRIMARY)) +
        theme_classic()
```

```{r, fig.asp=.8, fig.width=7.2, message=FALSE, warning=FALSE}
Europe <- c("AZORES (PORTUGAL)", "ALBANIA", "ANDORRA", "AUSTRIA", "BELARUS", "BELGIUM", "BOSNIA-HERZEGOVINA", "BOSNIA", "HERZEGOVINA2", "BULGARIA", "CROATIA", "CYPRUS", "CZECH REPUBLIC", "DENMARK", "FINLAND", "FRANCE", "GEORGIA", "GEROMANY", "GREECE", "HUNGARY", "ICELAND", "IRELAND", "ITALY", "KOSOVO", "LIECHTENSTEIN", "LUXEMBOURG", "MACEDONIA", "MALTA", "MOLDOVA", "MONACO", "MONTENEGRO", "NETHERLANDS", "NORWAY", "POLAND", "PORTUGAL", "ROMANIA", "SAN MARINO", "SERBIA", "SERBIA AND MONTENEGRO", "SLOVAKIA", "SLOVENIA", "SPAIN", "SWEDEN", "SWITZERLAND", "TURKEY", "UKRAINE", "UK", "UK TERRITORY", "VATICAN CITY")
quakes_EU <- NOAAearthquakes %>%
        filter(COUNTRY %in% Europe & YEAR >= 0)

# single country without labels
ggplot() +
        geom_timeline(data = quakes_EU, aes(x = date, y = COUNTRY, color = TOTAL_DEATHS, size = EQ_PRIMARY)) +
        theme_classic()

ggplot() +
        geom_timeline(data = quakes_EU, aes(x = date, y = COUNTRY, color = TOTAL_DEATHS, size = EQ_PRIMARY)) +
        geom_timeline_label(data = quakes_EU, aes(x = date,
                                                 y = COUNTRY,
                                                 magnitude = EQ_PRIMARY,
                                                 label = LOCATION_NAME,
                                                 n_max = 1)) +
        theme_classic()


```

```{r, message=FALSE, warning=FALSE, fig.width = 7.2, fig.asp = .5}
SouthAmerica <- c("ARGENTINA", "BOLIVIA", "BRAZIL", "CHILE", "COLOMBIA", "ECUADOR", "GUYANA", "PARAGUAY", "PERU", "SURINAME", "URUGUAY", "VENEZUELA")
quakes_SA <- NOAAearthquakes %>%
        filter(COUNTRY %in% SouthAmerica & YEAR >= 0)

# single country without labels
ggplot() +
        geom_timeline(data = quakes_SA, aes(x = date, y = COUNTRY, color = TOTAL_DEATHS, size = EQ_PRIMARY)) +
        theme_classic()


ggplot() +
        geom_timeline(data = quakes_SA, aes(x = date, y = COUNTRY, color = TOTAL_DEATHS, size = EQ_PRIMARY)) +
        geom_timeline_label(data = quakes_SA, aes(x = date,
                                                 y = COUNTRY,
                                                 magnitude = EQ_PRIMARY,
                                                 label = LOCATION_NAME,
                                                 n_max = 1)) +
        theme_classic()

```

#### Conclusion

As we have seen in the `README.md`, the timeline functions work very well for 1-3 countries. We see here, it works for multiple countries, too, however, the labeling should be improved on. We see that the labels are being cut off by the ggplot-object.  
Now let's see what the leaflet map shows us for these subsets

## Earth quakes and Leaflet map

We directly take the above defined subsets:  
- earth quakes in Central Europe  
- earth quakes in Europe  
- earth quakes in South America  

**(!!)** You'll see on the map that the countries are sourrended with a polygon. Please note that only countries which encountered and earth quake are marked with a polygon. If nothing happend, no polygon is shown.  

### Central Europe - A.D. until today
```{r, fig.width = 7.2, fig.asp = 1}
require(leaflet)

quakes_CentEU %>%
        dplyr::mutate(popup_text = eq_create_label(.)) %>%
        eq_map(annot_col = "popup_text")
```

### Europe - 2010 until today
```{r, fig.width = 7.2, fig.asp = 1}
quakes_EU %>%
        dplyr::filter(lubridate::year(date) >= 2010) %>% # additional subset, or else too many points
        dplyr::mutate(popup_text = eq_create_label(.)) %>%
        eq_map(annot_col = "popup_text")
```

### South America - 2010 until tolday
```{r, fig.width = 7.2, fig.asp = 1}
quakes_SA %>%
        dplyr::filter(lubridate::year(date) >= 2010) %>% # additional subset, or else too many points
        dplyr::mutate(popup_text = eq_create_label(.)) %>%
        eq_map(annot_col = "popup_text")
```

#### Conclusion

The leaflet maps work very well. Future work still can be done, for instance the circle representations of the earth quakes can be improved on (play around with transparency and size, etc.)
